Network Deliberation: The role of network structure in large-scale, internet-enabled, participatory decision-making

### 2023-02-16 UMich/front_materials/abstract.tex
Change Networked Deliberation to Network Deliberation.

As groups grow in size, they gain access to additional resources, creating opportunities for collective intelligence and collective action.
However, at very large scales, group decision-making becomes prohibitively slow and difficult to coordinate.
Traditional solutions include representative decision-making and/or a shift from deliberation to voting.
Both approaches sacrifice desirable properties.
Representative decision-making loses the potential benefits of collective intelligence and introduces hierarchies that may place the interests of specific individuals ahead of the interests of the group.
Voting sacrifices generativity: allowing a choice between predefined options, without allowing for improvement to those options.
Arrow’s impossibility theorem also fundamentally limits the fairness of voting.
This project proposes Networked Deliberation as a potential means of effective large-scale decision-making.
In Networked Deliberation, members of a large group are repeatedly partitioned into small deliberative pods.
Overlap between pods at different stages enables group-wide diffusion of information and preferences.
Different methods for assigning members to pods result in different network topologies.
This work combines observational study, agent-based modeling, and an experiment to evaluate and better understand network deliberation.
An empirical observation of WikiProjects on the English-language Wikipedia identifies the network properties of the most effective projects.
A simple agent-based model of Network Deliberation shows improvements over conventional deliberation in the presence of strong social influence.
Finally, a controlled experiment studies network deliberation in a real world setting, tracking how individual preferences evolve, and finding evidence that network deliberation provides protection against negative consequences of social influence.
This work seeks to provide a framework that enables an understanding of
how very large groups can quickly develop a consensus, enabling a more effective use of shared resources to achieve common goals.

>>>>

As groups grow in size, they gain access to additional resources, creating opportunities for collective intelligence and collective action.
However, at very large scales, group decision-making becomes prohibitively slow and difficult to coordinate.
Traditional solutions include representative decision-making and/or a shift from deliberation to voting.
Both approaches sacrifice desirable properties.
Representative decision-making loses the potential benefits of collective intelligence and introduces hierarchies that may place the interests of specific individuals ahead of the interests of the group.
Voting sacrifices generativity: allowing a choice between predefined options, without allowing for improvement to those options.
Arrow’s impossibility theorem also fundamentally limits the fairness of voting.
This project proposes Network Deliberation as a potential means of effective large-scale decision-making.
In Network Deliberation, members of a large group are repeatedly partitioned into small deliberative pods.
Overlap between pods at different stages enables group-wide diffusion of information and preferences.
Different methods for assigning members to pods result in different network topologies.
This work combines observational study, agent-based modeling, and an experiment to evaluate and better understand network deliberation.
An empirical observation of WikiProjects on the English-language Wikipedia identifies the network properties of the most effective projects.
A simple agent-based model of Network Deliberation shows improvements over conventional deliberation in the presence of strong social influence.
Finally, a controlled experiment studies network deliberation in a real world setting, tracking how individual preferences evolve, and finding evidence that network deliberation provides protection against negative consequences of social influence.
This work seeks to provide a framework that enables an understanding of
how very large groups can quickly develop a consensus, enabling a more effective use of shared resources to achieve common goals.
